---
title: "ToOoOlTiPs: An R package for Customizable Tooltips in Interactive Graphics"
date: "`r Sys.Date()`"
abstract: >
  An abstract of less than 150 words.
draft: true
author:  
  # see ?rjournal_article for more information
  - name: Quietest Quokka
    affiliation: University of Little Mates
    address:
    - Department of Letter Q
    - Somewhere, Australia
    url: https://www.britannica.com/animal/quokka
    orcid: 0000-1721-1511-1101
    email:  qquo@ulm.edu
  - name: Bounciest Bilby
    url: https://www.britannica.com/animal/bilby
    email: bbil@ulm.edu
    orcid: 0000-0002-0912-0225
    affiliation: University of Little Mates
    address:
    - Department of Letter B
    - Somewhere, Australia
creative_commons: CC BY
output: 
  - rjtools::rjournal_pdf_article
  - rjtools::rjournal_web_article
bibliography: paper-with-errors.bib

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(plotly)
library(ggplot2)
library(palmerpenguins)
```

# Introduction

Interactive graphics is a type of visualization that allows users to interactively inspect a plot. One of the most basic interactivity is through tooltips, where users can request additional information about a plot element by mouse hovering ...

This paper will first review some R packages on interactive graphics and their tooltip implementations. Next, a new package \CRANpkg{ToOoOlTiPs} will be proposed for customizing tooltips. Some gallery plots will then be given to showcase how these tooltips help users to better read the graphics. 

# Literature review

Some packages on interactive graphics include \CRANpkg{plotly} [@plotly] that interfaces with Javascript for web-based interactive graphics, \CRANpkg{crosstalk} [@crosstalk] that specializes cross-linking elements across individual graphics, and most recently, \CRANpkg{tsibbletalk} [@RJ-2021-050] for working with time series data in a tsibble structure ...

# Customizing tooltip design with \pkg{ToOoOlTiPs}

\pkg{ToOoOlTiPs} is a packages for customizing tooltips in interactive graphics, it features ...

# A gallery of tooltips

We first show a baseline plot, in Figure \ref{fig:penguins-plotly}, made by the \pkg{plotly} package with \CRANpkg{palmerpenguins} data [@palmerpenguins]. 

```{r penguins-plotly, echo = TRUE, fig.height = 5, fig.cap="A basic interactive plot made with the plotly package on palmer penguin data. Three species of penguins are plotted with bill depth on the x-axis and bill length on the y-axis. When hovering on a point, a tooltip will show the exact value of the bill depth and length for that point, along with the species name."}
p <- penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, 
             color = species)) + 
  geom_point()
ggplotly(p)
```

Now we will re-create the same plot using the \pkg{ToOoOlTiPs} package for different tooltips designs ...

# Summary

We have displayed various tooltips that are available in the package \pkg{ToOoOlTiPs}. These tooltips ...
