---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rjtools

<!-- badges: start -->
<!-- badges: end -->

The goal of rjtools is to make it easier for AUTHORS of R Journal
articles to 

1. create a new paper from a template that will produce both html and pdf versions in the format needed by the journal, and 
2. to run various checks on their paper, that it satisfies various requirements, in preparation for submitting it to the journal.

## Installation

<!-- You can install the released version of rjtools from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("rjtools") -->
<!-- ``` -->

<!-- And the development version from [GitHub](https://github.com/rjournal/rjtools) with: -->

You can install the development version from [GitHub](https://github.com/rjournal/rjtools) with: 

``` r
# install.packages("remotes")
remotes::install_github("rjournal/rjtools")
```

## Getting started with a new article

To create an article template under the project directory: 

To create an article template under the project directory:

```{r eval=FALSE, echo=TRUE}
library(rjtools)
create_article(file_name = "my_article")
```

## Checking that your paper satisfies various conditions

Explore the various `check_XXX` functions that will check the format of
your article, and some submission requirements.
