---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rjtools

<!-- badges: start -->
<!-- badges: end -->

The goal of rjtools is to make it easier for authors of R Journal articles to start a new paper, and to check their paper is in suitable condition to submit.

## Installation

<!-- You can install the released version of rjtools from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("rjtools") -->
<!-- ``` -->

<!-- And the development version from [GitHub](https://github.com/) with: -->

You can install the development version from [GitHub](https://github.com/) with: 

``` r
# install.packages("remotes")
remotes::install_github("rjournal/rjdistill")
remotes::install_github("rjournal/rjtools")
```
## Example

To create an article template under the project directory: 

```{r example, eval = FALSE}
library(rjtools)
create_article(file_name = "my_article")
```

## Check functions

Explore the various `check_XXX` functions that will check the format of your article, and some submission requirements.
