---
title: "Check your Article Before Submission"
author: Sherry Zhang
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Check}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(rjtools)
```

In the `inst/paper-with-errors` folder is an example article that has both html and pdf version rendered and it will be used to demonstrate the check functions in the `rjtools`. Notice that the check function implemented here is only an initial automated check and we recommend authors to also read [instructions for Authors](https://journal.r-project.org/share/author-guide.pdf) before submission. 

There are 9 functions in total for authors to check against and these checks can be completed in one go with `initial_check_article()`, given a path to the article folder: 

```{r eval = FALSE}
initial_check_article(path = here::here("paper-with-errors"))
```

The check result will be saved in an initial_check.log file, as well as displayed in the console. Checking on each item is also supported by individual `check_*()` function: 

  - `check_wrappers()`
  - `check_filenames()`
  - `check_unnecessary_files()`
  - `check_cover_letter()`
  
  - `check_title()`
  - `check_section()`
  - `check_spelling()`
  - `check_proposed_pkg()`
  - `check_packages_available()`

Below we will apply each check function on the example article we create

## Wrappers

```{r}
check_wrappers(path = here::here("inst/paper-with-errors"))
```

A properly rendered article should have both the `RJwrapper.tex` and `RJwrapper.pdf` file and our article passes this check. Good!

## Filenames

```{r}
check_filenames(path = here::here("inst/paper-with-errors"))
```

The .bib file is named `skeleton.bib` when created and we recommend setting a consistent name across the .bib, .tex, and .Rmd file.  A fix to this is to rename the `skeleton.bib` as the file name you used when create the article, in our case, `paper-with-errors`. Also remember to make the same change in the Rmd YAML, so it can find the .bib file during knitting!

## Presence of unnecessary files

```{r} 
check_unnecessary_files(path = here::here("inst/paper-with-errors")) 
```

This function checks that the folder doesn't contain the `RJtemplate.tex` file, which is icluded the previous R journal template. The [authors instruction](https://journal.r-project.org/share/author-guide.pdf) requires the authors to replace the content of this `RJtemplate.tex` file with article content and name with the lowercase concatenated author name, so it is essentially the `.tex` file automatically generated in the previous section. Our article don't have unnecessary file here. Good :)

## Presence of a motivating letter

```{r} 
check_cover_letter(path = here::here("inst/paper-with-errors")) 
```

Cover letter should be provided for article of type "add-on package" and saved at the same folder as the article. The function checks whether there is a file has the word "motivation" in its name and the note says our sample article has not yet have a motivational letter. Bad!

## Article title is in title case

```{r} 
check_title(path = here::here("inst/paper-with-errors")) 
```

The article title needs to be in the title case and the check is not happy with our title here. Use `tools::toTitleCase()` to get the correct title case format:

```{r}
tools::toTitleCase("ToOoOlTiPs: An R package for Customizable Tooltips in Interactive Graphics")
```

Here we forget to capitalise "P" in the word "package" - better make this change before submission! Also if the title contains the package you're proposing, i.e `ToOoOlTiPs` in our article, use the `\pkg{.}` tag rather than `\CRANpkg{.}`.

## Section title is in sentence case

```{r} 
check_section(path = here::here("inst/paper-with-errors")) 
```

Section titles need to be in the sentence case and we are all good!

## Spelling

```{r} 
check_spelling(path = here::here("inst/paper-with-errors")) 
```

The spelling check is performed from the abstract to bibliography section, excluding the comments in the code chunks. Only words with all lowercase letters are checked to avoid picking up on acronyms and names.
The underlying checking is done through the `hunspell` package with US spelling by default. British spelling can be checked with argument `dic = "en_GB"`. The R journal welcomes both the U.S and British spelling, along with other national variation, as long as consistency is maintained. Notice that the spelling check may pick up some domain specific vocabularies and authors may want to further confirm if the suggested spelling mistakes are true mis-spellings. In our article, "tooltip" is a specialised vocabulary in interactive graphics, which is not really a mis-spelling, so we are free of spelling mistake!


## Package is available on CRAN

```{r eval = FALSE} 
check_proposed_pkg() 
```

The package proposed in an article should already be available on CRAN before article submission. `check_proposed_pkg` prompts the following in the console:

  > What's the name of package being proposed in the article? If none, please enter 0. 

and authors need to enter the exact package name for an availability check on CRAN download with the `cranlogs` package. For our article, enter "ToOoOlTiPs" in the console and the check would say 

  > NOTE: No CRAN activities detected for package ToOoOlTiPs
  
Definitely CRAN can't find our package since it is a hypothetical one!  

## Package availability

```{r eval = FALSE} 
check_packages_available(path = here::here("inst/paper-with-errors")) 
```

All the packages mentioned with a `\CRANpkg{.}` or `\BIOpkg{.}` tag need to be available on CRAN or Bioconductor and `check_packages_available` makes sure this is the case. For our article, we get 

 > ! ERROR: 1 of 5 package(s) not available on CRAN: ToOoOlTiPs
 
which is not really an issue :)






