---
title: "Create an R Journal Article"
author: Sherry Zhang
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{create_article}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(rjtools)
```

With the advances in html delivery of research articles through the web, the R Journal now has a new format of it's website using a version of the distill package. This will allow incorporation of interactive graphics directly into articles, and easier reading of the research online. 

The Rmd template in `rjtools` creates an article in the format needed for the R journal distill web site. The `rjtools` packages is provided to help authors meet these requirements, with an intention that it would help authors to write an R journal article, a little like the way that `devtools` help developers build an R package. In this vignette, I will walk you through how to use `rjtools` to create an R journal article using the build-in template. 

## Create an Article

The `create_article()` function will create these files and folder: 

  - **figures/**: folder for saved figures or images
  - **RJwrapper.tex**: the tex template for compile an R journal article
  - **Rjournal.sty**: the R journal style sheet
  - **article.Rmd**: the main rmd document
  - **skeleton.bib**: a bibtext file for adding reference
  - **biostats.csl**: the R journal reference style
  - **motivation-letter.md**: a motivation letter template for type add-on package

The `.sty` and `.csl` files make sure all the R Journal articles follow the same LATEX and reference style and should not be modified unless there's a strong reason to do so. The `motivation-letter.md` can be deleted if the article type is not an add-on package.

The `file_name` argument can be used to change the name of the main rmd document and we recommend using the concatenation of all the authors' family name. For example, an article by Quietest Quokka and Bounciest Bilby can be created by

```{r eval = FALSE}
create_article(file_name = "quokka-bilby")
```

In this case, the `article.Rmd` will become `quokka-bilby.Rmd`. If you have specified the file name as we recommended, be sure to also rename the `.bib` file and modify the bibliography reference in the `quokka-bilby.Rmd` YAML. R journal requires a consistent name between the bib and article file, otherwise check function `check_filenames` will pick it up. 

## Your first knit

The `rjtools::rjournal_web_article` output specified in the YAML will create the article in both html and pdf format. Additional files created during the knit are (assuming file name of `quokka-bilby`):

  - **quokka-bilby.html**: the html article 
  - **RJwrapper.pdf**: the pdf article 
  - **quokka-bilby.tex**: the LaTex file 
  - **quokka-bilby.R**: an R file that contains all the code 
  
## Note: `rticles` template

The `rticles` package contains a template for creating an R Journal article, which you might have used before if you have been already working with Rmarkdown for writing papers. This template creates an article that will be compiled with latex into a pdf format paper. It is using the current style file. If this new distill version of the delivery of the R Journal is successful, we will also synchronise the template the one in `rjtools`. 
