---
title: "Create an R Journal Article"
author: Sherry Zhang
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{create_article}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(rjtools)
```

With the advances in html delivery of research articles through the web, the R Journal now has a new format of it's website using a version of the distill package. This will allow incorporation of interactive graphics directly into articles, and easier reading of the research online. 

The Rmd template in `rjtools` creates an article in the format needed for the R journal distill web site. The `rjtools` packages is provided to help authors meet these requirements, with an intention that it would help authors to write an R journal article, a little like the way that `devtools` help developers build an R package. In this vignette, I will walk you through how to use `rjtools` to create an R journal article using the build-in template. 

## Create an Article

The package contains a template for the R journal articles. The `create_article()` function will create a paper folder in the root directory with three files: 

  - article.Rmd, 
  - skeleton.bib, 
  - RJwrapper.tex, and 
  - Rjournal.sty

The folder name and bib/Rmd file name can be changed with argument `dir_path` and `file_name`. The R journal requires the .bib file and the .Rmd file to have the same name ("article" as created by default). We recommend using a concatenation of the family names of all the authors as a file name. 

For example, an article by Quietest Quokka and Bounciest Bilby can be created by

```{r eval = FALSE}
create_article(file_name = "quokka-bilby")
```


This will generate 
  
  - quokka-bilby.Rmd, 
  - skeleton.bib, 
  - RJwrapper.tex, and 
  - Rjournal.sty
  
in a folder called "paper". We recommend a consistent name between the bib and article file, so you may want to decide to rename `skeleton.bib` to `quokka-bilby.bib` and also change the bibliography reference in the `quokka-bilby.Rmd` YAML.

## Note: `rticles` template

The `rticles` package contains a template for creating an R Journal article, which you might have used before if you have been already working with Rmarkdown for writing papers. This template creates an article that will be compiled with latex into a pdf format paper. It is using the current style file. If this new distill version of the delivery of the R Journal is successful, we will also synchronise the template the one in `rjtools`. 
