---
title: "Create an R Journal Article"
author: Sherry Zhang
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{create_article}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  comment = "#>"
)
```

```{r setup, echo = FALSE}
library(rjtools)
```

R journal features an automated online review process that requires certain structure and format on the submitted articles. To help authors meet these requirements, we create the `rjtools` package, with an intention that it would help authors to write an R journal article, just as how `devtools` would help developers build a package. In this vignette, I will walk you through how to use `rjtools` to create an R journal article using the build-in template. 

## Create an Article

The package contains a template for the R journal articles. The `create_article()` function will create a paper folder in the root directory with three files: 

  - article.bib, 
  - article.Rmd, and 
  - RJwrapper.tex

The folder name and bib/Rmd file name can be changed with argument `dir_path` and `file_name`. Notice that the R journal requires the .bib file and the .Rmd file to have the same name ("article" as created by default) and we recommend connecting the surname of all the authors as a name. 

For example, an article by Di Cook, Catherine Hurley, Simon Urbanek, Mike Kane and Gavin Simpson can be created by

```{r eval = FALSE}
create_article(file_name = "cook-hurley-urbanek-kane-simpson")
```


This will generate 
  
  - cook-hurley-urbanek-kane-simpson.bib, 
  - cook-hurley-urbanek-kane-simpson.Rmd, and 
  - RJwrapper.tex
  
in a folder called "paper". One last step before start on writing is to make a change on the bibliography reference in the Rmd YAML. By default, the template generates `bibliography: RJreferences.bib` in the YAML and you want to replace `RJreferences` with the file name you just specified: `cook-hurley-urbanek-kane-simpson`.

