---
title: "create an article using `rjtools`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{create_article}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rjtools)
```

# Create an article 

To create a new R journal article, use the `create_article()` function in the `rjtools` package. The function will create a paper folder in the root directory with three files: article.bib, article.Rmd, and RJwrapper.tex by default. 

The name of the folder and bib/Rmd file can be customised by using the `dir_path` and `file_name` argument. R journal requires the bib file and the rmarkdown file to have the same name and we recommend them to be named as the surname of all the authors. 

For example, to create an article by Di Cook, Catherine Hurley, Simon Urbanek, and Michael Kane, use 

```{r eval = FALSE}
create_article(file_name = "cook-hurley-urbanek-kane")
```


This will generate "cook-hurley-urbanek-kane.Rmd", "cook-hurley-urbanek-kane.bib", and "RJwrapper.tex" under the paper folder. 

If you made change of the file name with `file_name`, please also change the reference to the bib file specified in the YAML of the rmarkdown file to the same name.


