---
title: "ToOoOlTiPs: An R Package for Customizable Tooltips in Interactive Graphics"
abstract: >
  An abstract of less than 150 words.
draft: true
author:  
  # see ?rjournal_article for more information
  - name: Quietest Quokka
    affiliation: University of Little Mates
    address:
    - Department of Letter Q
    - Somewhere, Australia
    url: https://www.britannica.com/animal/quokka
    orcid: 0000-1721-1511-1101
    email:  qquo@ulm.edu
  - name: Bounciest Bilby
    url: https://www.britannica.com/animal/bilby
    email: bbil@ulm.edu
    orcid: 0000-0002-0912-0225
    affiliation: University of Little Mates
    address:
    - Department of Letter B
    - Somewhere, Australia
date: "`r Sys.Date()`"
creative_commons: CC BY
output: 
  rjtools::rjournal_web_article
bibliography: skeleton.bib

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(plotly)
library(ggplot2)
library(palmerpenguins)
```

# Introduction

Interactive data graphics provides plots that allow users to interact them. One of the most basic types of interaction is through tooltips, where users are provided additional information about elements in the plot by moving the cursor over the plot.

This paper will first review some R packages on interactive graphics and their tooltip implementations. A new package \CRANpkg{ToOoOlTiPs} that provides customized tooltips for plot, is introduced. Some example plots will then be given to showcase how these tooltips help users to better read the graphics. 

# Background

Some packages on interactive graphics include \CRANpkg{plotly} [@plotly] that interfaces with Javascript for web-based interactive graphics, \CRANpkg{crosstalk} [@crosstalk] that specializes cross-linking elements across individual graphics.  The recent R Journal paper \CRANpkg{tsibbletalk} [@RJ-2021-050] provides a good example of including interactive graphics into an article for the journal. It has both a set of linked plots, and also an animated gif example, illustrating linking between time series plots and feature summaries.

# Customizing tooltip design with \pkg{ToOoOlTiPs}

\pkg{ToOoOlTiPs} is a packages for customizing tooltips in interactive graphics, it features these possibilities.

# A gallery of tooltips examples

Figure \ref{fig:penguins-plotly} shows an interactive plot of the \CRANpkg{palmerpenguins} data [@palmerpenguins], made using the \pkg{plotly} package. 

```{r penguins-plotly, echo = TRUE, fig.height = 5, fig.cap="A basic interactive plot made with the plotly package on palmer penguin data. Three species of penguins are plotted with bill depth on the x-axis and bill length on the y-axis. When hovering on a point, a tooltip will show the exact value of the bill depth and length for that point, along with the species name."}
p <- penguins %>% 
  ggplot(aes(x = bill_depth_mm, y = bill_length_mm, 
             color = species)) + 
  geom_point()
ggplotly(p)
```

# Summary

We have displayed various tooltips that are available in the package \pkg{ToOoOlTiPs}. 
