---
title: "Xyz: a greate package that I created"
abstract: >
  An abstract of less than 150 words.
draft: true
author:  
  # see ?rjournal_article for more information
  - name: Author One
    affiliation: Affiliation
    address:
    - line 1
    - line 2
    url: https://journal.r-project.org
    orcid: 0000-0002-9079-593X
    email:  author1@work
  - name: Author Two
    url: https://journal.r-project.org
    email: author2@work
    orcid: 0000-0002-9079-593X
    affiliation: Affiliation 1
    address:
    - line 1 affiliation 1
    - line 2 affiliation 1
    affiliation2: Affiliation 2
    address2:
    - line 1 affiliation 2
    - line 2 affiliation 2
date: "`r Sys.Date()`"
creative_commons: CC BY
output: 
  - rjtools::rjournal_web_article
  - rjtools::rjournal_pdf_article
# per R journal requirement, the bib filename should be the same as the output 
# tex file. Don't forget to rename the bib file and change this example value.
bibliography: sample-article.bib

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(plotly)
```

# Introduction

An introduction to the paper ...

# Another section (... that needs to be in the sentence case)

There will likely be several sections, including literature review with some references in the `.bib` file, some user case demos with code snippets as follows: 

```{r eval = FALSE, echo = TRUE}
x <- 1:10
plot(x)
```

and a plot with cross-reference and captions as in Figure \@ref(fig:myplot),

```{r myplot, fig.cap = "A caption should include three elements: 1) a one sentence summary of what the plot is about; 2) the details of elements being plotted; and 3) the messages from the plot."}
x <- 1:10
plot(x)
```

## Summary

This file is only a basic article template. For full details of _The R Journal_ style and information on how to prepare your article for submission, see the [Instructions for Authors](https://journal.r-project.org/share/author-guide.pdf).
